---
title: "Lattice Multivariate Data in R"
output: html_document
---

### Chapter 1
```{r}
# install.packages('mlmRev')
library('lattice')
library('mlmRev')
data(Chem97, package = 'mlmRev')
xtabs(~ score, data=Chem97)
```


- Is the gcsescore different for different values of score? 
We can produce a histogram of gcsescore for each score, placing them all together on a single page: 

```{r}
library('lattice')
histogram(~ gcsescore | factor(score), data = Chem97)
```

#### The Trellis Call

```{r}
densityplot(~ gcsescore | factor(score), data = Chem97, plot.points = FALSE, ref = TRUE)
```


#### Superposition
```{r}
densityplot(~ gcsescore, data = Chem97, groups = score,
plot.points = FALSE, ref = TRUE,
auto.key = list(columns = 3))
```

```{r}
tp1 <- histogram(~ gcsescore | factor(score), data = Chem97)
tp2 <-
densityplot(~ gcsescore, data = Chem97, groups = score,
plot.points = FALSE,
auto.key = list(space = "right", title = "score"))

class(tp1)
class(tp2)

summary(tp1)
summary(tp2)

plot(tp1)
plot(tp2)

```

Arranging multiple trellis plots
```{r}
plot(tp1, split = c(1, 1, 1, 2))
plot(tp2, split = c(1,2,1,2), newpage = FALSE)
```



### Chapter 3 - Visualizing Univariate Distributions

```{r}
densityplot(~ eruptions, data = faithful)
```

```{r}
densityplot(~ eruptions, data=faithful, kernel='rect', bw=0.2, plot.points='rug', n=200)
```


```{r}
library("latticeExtra")
data(gvhd10)
densityplot(~log(FSC.H) | Days, data = gvhd10,
plot.points = FALSE, ref = TRUE, layout = c(2, 4))
```


```{r}
histogram(~log2(FSC.H) | Days, gvhd10, xlab = "log Forward Scatter", type = "density", nint = 50, layout = c(2, 4))
```


Normal QQ plots of average GCSE score for different final scores in A level chemistry.
```{r}
qqmath(~ gcsescore | factor(score), data = Chem97, f.value = ppoints(100))
```


```{r}

qqmath(~ gcsescore | gender, Chem97, groups = score, aspect = "xy", f.value = ppoints(100), auto.key = list(space = "right"), xlab = "Standard Normal Quantiles",
ylab = "Average GCSE Score")

```


Box and Whisker Plots
```{r}
bwplot(factor(score) ~ gcsescore | gender, data = Chem97, xlab = "Average GCSE Score")

# Or alternatively by changing angle..
bwplot(gcsescore^2.34 ~ gender | factor(score), data = Chem97, varwidth = TRUE, layout = c(6, 1),
         ylab = "Transformed GCSE score")


bwplot(Days ~ log(FSC.H), data = gvhd10,
xlab = "log(Forward Scatter)", ylab = "Days Past Transplant")


```


Strip Plots
```{r}

stripplot(factor(mag) ~ depth, quakes)


stripplot(depth ~ factor(mag), quakes,
            jitter.data = TRUE, alpha = 0.6,
xlab = "Magnitude (Richter)", ylab = "Depth (km)")
```


Cleveland dot plot
```{r}
VADeaths

dotplot(VADeaths, groups = FALSE)

dotplot(VADeaths, groups = FALSE,
          layout = c(1, 4), aspect = 0.7,
          origin = 0, type = c("p", "h"),
          main = "Death Rates in Virginia - 1940",
          xlab = "Rate (per 1000)")


dotplot(VADeaths, type = "o",
auto.key = list(lines = TRUE, space = "right"), main = "Death Rates in Virginia - 1940",
xlab = "Rate (per 1000)")
```


Barcharts
```{r}
barchart(VADeaths, groups = FALSE,
layout = c(1, 4), aspect = 0.7, reference = FALSE, main = "Death Rates in Virginia - 1940",
xlab = "Rate (per 100)")


barchart(prop.table(postdoc, margin = 1), xlab = "Proportion", auto.key = list(adj = 1))


dotplot(prop.table(postdoc, margin = 1), groups = FALSE, xlab = "Proportion",
par.strip.text = list(abbreviate = TRUE, minlength = 10))

dotplot(prop.table(postdoc, margin = 1), groups = FALSE, index.cond = function(x, y) median(x),
xlab = "Proportion", layout = c(1, 5), aspect = 0.6, scales = list(y = list(relation = "free", rot = 0)), prepanel = function(x, y) {
              list(ylim = levels(reorder(y, x)))
          },
          panel = function(x, y, ...) {
              panel.dotplot(x, reorder(y, x), ...)
})
```


Barcharts and discrete distributions
```{r}
gcsescore.tab <- xtabs(~gcsescore + gender, Chem97)
gcsescore.df <- as.data.frame(gcsescore.tab)
gcsescore.df$gcsescore <-
as.numeric(as.character(gcsescore.df$gcsescore))
xyplot(Freq ~ gcsescore | gender, data = gcsescore.df,
type = "h", layout = c(1, 2), xlab = "Average GCSE Score")
```



```{r}
```


Visualizing Categorical Data
```{r}
score.tab <- xtabs(~score + gender, Chem97)
score.df <- as.data.frame(score.tab)
barchart(Freq ~ score | gender, score.df, origin = 0)
```



```{r}
```



```{r}
```



```{r}
```



```{r}
```



```{r}
```