---
title: "SVM Example with Iris Data in R"
output: html_document
---

Copied notes from: http://rischanlab.github.io/SVM.html

```{r}
library('e1071')
```

```{r}
head(iris,5)
dim(iris)
attach(iris)
str(iris)
```

Divide Iris data to x (the variables) and y the classes
```{r}
x = subset(iris, select=-Species)
y = Species
```

Create SVM Model and show summary
```{r}
svmModel = svm(x,y, data=iris)
summary(svmModel)
```


Run prediction ...
```{r}
pred = predict(svmModel, x)
```


See the confusion matrix result of prediction
```{r}
table(pred,y)
```

Tuning SVM to find the best cost and gamma:

```{r}
svmTune = tune(svm, train.x = x, train.y = y, kernel='radial', ranges=list(10^(-1:2), gamma=c(0.5, 1,2)))
print(svmTune)
```

After finding the best cost and gamma, we can create svm model again:

```{r}
svmModelPostTune = svm(x,y,data=iris, kernel='radial', cost=0.1, gamma=0.5)
summary(svmModelPostTune)
```


Re-running the prediction with the new tuned model:

```{r}
pred = predict(svmModelPostTune,x)
```

See the new tuned model confusion matrix:

```{r}
table(pred,y)
```
